Recruitment rate estimation in waterfowl using a non-invasive method

### 2021-12-16 Code/Draft_ruddy_duck_recruitment_rate.Rmd
readme edit

The simple relationship (a bit more complex if a species with delayed maturity is considered [see @Robertson2008]) between population growth rate, adult survival, and recruitment rate can be used to test the accuracy of the order of magnitude of recruitment rates estimated from the counting and the sampling methods, see Equation \@ref(eq:GR). This relationship comes from a simple reasoning for a closed population: the population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e. the recruitment in year $t$. The growth rate of a population is thus the sum of the adult survival rate and the recruitment rate [@Flint2015].

@@

From the evolution of the breeding population size, population growth rate can be assessed every year to catch its variability range \@ref(eq:GR), or can be smoothed over a consistent time period to provide average population growth rate, see Equation \@ref(eq:L), which is a linear regression on the logarithm scale.

@@

The recruitment rates from the counting method are in the same range for the two populations. The higher variability observed in France is not unexpected because the French time series is longer and covers a larger harvest spectrum than in UK. There is no temporal correlation between the two populations, so it suggests that the recruitment rate is more likely influenced by regional-scale factors than large-scale ones. This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of lay, which is tightly related to local weather parameters, i.e. the spring temperature, the cumulative rainfall, and the water levels [@Folliot2017; @Blums2004; @Dzus1998].

Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation to the outputs of the counting method demonstrates its ability to track the recruitment rate variability. This result is quite robust because the correlation is obvious with only a five-year time series. However, this conclusion might be overrated because two successive population size values are part of the recruitment rate, which mechanically scale the outputs. It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate (Figure \@ref(fig:sup)A).

A significant change in adult sex ratio would bias the estimation of recruitment rate. However, since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong. Significant variations can then occur only every multiple years, especially if it is a long-living species. Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.

The counting method always provides recruitment rates significantly lower than the sampling method, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl. This is a robust result because the approaches used to control ruddy ducks differ strongly from usual hunting practices and are apriori less selective. It is especially clear in the UK, for which control operations occur in winter using long-range rifles that prevents the bird from escaping from the threat. The assumption on the accuracy of the counting method is then not excluded from the consistency of this result.

*survival consistency*

Annual adult survival rates deduced from observed population growth rates and recruitment estimates are on average in line with the literature **ref required**.

@@

>>>>

The simple relationship (a bit more complex if a species with delayed maturity is considered, see @Robertson2008) between population growth rate, adult survival, and recruitment rate can be used to test the accuracy of the order of magnitude of recruitment rates estimated from the counting and the sampling methods, see Equation \@ref(eq:GR). This relationship comes from a simple reasoning for a closed population: the population size in year $t$ is equal to the number of breeders that survived over year $t - 1$ plus the offspring produced in year $t - 1$ that survived until the reproduction period of year $t$, i.e. the recruitment in year $t$. The growth rate of a population is thus the sum of the adult survival rate and the recruitment rate [@Flint2015].

@@

From the evolution of the breeding population size, population growth rate can be assessed every year to catch the variability range \@ref(eq:GR), or can be smoothed over a consistent time period to provide average population growth rate, see Equation \@ref(eq:L), which is a linear regression on the logarithm scale.

@@

The recruitment rates from the counting method are in the same range for the two populations, which demonstrate a certain consistency in the results obtained from the method. The higher variability observed in France is not unexpected because the French time series is longer and covers a larger harvest spectrum than in UK. There is no temporal correlation between the two populations, so it suggests that the recruitment rate fluctuations are more likely influenced by regional-scale factors than large-scale ones. This is consistent with the literature because both nesting/hatching success and juvenile survival are conditional to the onset of lay, which is tightly related to local weather parameters, e.g. the spring temperature, the cumulative rainfall, and the water levels [@Folliot2017; @Blums2004; @Dzus1998].

Assuming that the sample data provide a good picture of the interannual variability of the recruitment rate, the good correlation to the outputs of the counting method demonstrates the ability of this method to track the recruitment rate variability. This result is quite robust because the correlation is obvious with only a five-year time series. However, this conclusion might be overrated because two successive population size values are part of the recruitment rate, which mechanically scale the outputs. It explains why the correlation on the immature proportion, the other component of the recruitment rate, is a bit more noisy than for the recruitment rate (Figure \@ref(fig:sup)A). The uncertainty of the estimates is very close to the outputs from the sampling method. Even though the counting method is more indirect to infer on the recruitment rate, it performs to provide a good confidence of the outputs which is satisfying. A significant change in adult sex ratio would bias the evolution of the recruitment rate. However, since this ratio integrates many age-cohorts, the temporal autocorrelation is structurally strong. Significant variations can then occur only every multiple years, especially if it is a long-living species. Consequently, it is not necessary to monitor and update the adult sex ratio on a yearly base.  
The counting method always provides recruitment rates significantly lower than the sampling method by factor close to 2 on average, a result that is in line with the hypothesis that the harvest is generally selective towards immature individuals in waterfowl because they are more vulnerable to hunters [@Bellrose1980; @Fox2014]. This is a robust result because the approaches used to control ruddy ducks differ strongly from usual hunting practices and are apriori less selective. It is especially clear in the UK, for which control operations occur in winter using long-range rifles that prevents the bird from escaping from the threat. The assumption on the accuracy of the counting method is then not excluded from the consistency of this result.

Annual adult survival rates deduced from observed population growth rates and recruitment estimates are on average in line with the literature **ref required**. which confirms that the counting method is poorly biased if there is some bias.

@@

The harvest pressure, simplified into a two-category variable, explains already 19.3% and 4.6% of the variance of the recruitment rate and the adult survival rate of the French population. The harvest pressure described as a continuous variable is expected to explain even more the recruitment rate variability. The harvest pressure applied in France is thus a factor that likely induces a strong immature mortality, but a moderate one on the adults.
